---
title: "Monthly Retail Sales Trend Analysis"
author: "Sayooj Sathiyan T"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
---

**Enrollment No.: 2505101270260**

# 1. Introduction

This Tiny Project analyses monthly retail sales data using R. The study examines
sales, advertising expenditure, customer count, average order value and returns.
It demonstrates data importing, preprocessing and cleaning, statistical
exploration, graph creation, interpretation, findings and conclusion.

# 2. Objectives

- Import monthly sales data from a CSV file.
- Check the structure and quality of the data.
- Perform basic preprocessing and cleaning.
- Calculate descriptive statistics.
- Study relationships between sales, advertising and customers.
- Create graphs and interpret the results.
- Summarize findings and conclusion.

# 3. Database / Data Importing

```{r import}
sales_data <- read.csv("monthly_sales_data.csv")
sales_data
```

**Explanation:** `read.csv()` reads the CSV file into an R data frame.

# 4. Data Preprocessing and Cleaning

```{r cleaning}
str(sales_data)
colSums(is.na(sales_data))
sum(duplicated(sales_data))

sales_data$Month <- as.character(sales_data$Month)
sales_data <- sales_data[!duplicated(sales_data), ]

summary(sales_data)
```

**Explanation:** `str()` checks the structure and data types. `is.na()` checks for
missing values, while `duplicated()` checks duplicate records. The cleaning step
removes any duplicate rows before analysis.

# 5. Data Analysis and Statistical Exploration

## 5.1 Descriptive statistics

```{r descriptive}
mean_sales <- mean(sales_data$Sales)
median_sales <- median(sales_data$Sales)
sd_sales <- sd(sales_data$Sales)
mean_customers <- mean(sales_data$Customers)

mean_sales
median_sales
sd_sales
mean_customers
```

**Explanation:** `mean()` calculates the average, `median()` gives the middle
observation, and `sd()` measures the spread of sales values.

## 5.2 Total sales and returns

```{r totals}
total_sales <- sum(sales_data$Sales)
total_returns <- sum(sales_data$Returns)

total_sales
total_returns
```

**Explanation:** `sum()` adds all monthly observations to obtain the total sales
and total returns for the study period.

## 5.3 Correlation analysis

```{r correlation}
cor_advertising_sales <- cor(
  sales_data$Advertising_Spend,
  sales_data$Sales
)

cor_customer_sales <- cor(
  sales_data$Customers,
  sales_data$Sales
)

cor_advertising_sales
cor_customer_sales
```

**Explanation:** `cor()` measures the strength and direction of a linear
relationship between two numeric variables.

# 6. Graph and Plot Creation

## 6.1 Monthly sales trend

```{r sales-trend}
plot(
  sales_data$Month,
  sales_data$Sales,
  type = "o",
  main = "Monthly Sales Trend",
  xlab = "Month",
  ylab = "Sales",
  las = 2
)
```

**Interpretation:** The line graph shows how sales changed over the 24-month period.

## 6.2 Monthly customer count

```{r customers}
barplot(
  sales_data$Customers,
  names.arg = sales_data$Month,
  main = "Monthly Customer Count",
  xlab = "Month",
  ylab = "Customers",
  las = 2
)
```

**Interpretation:** The bar chart compares the number of customers month by month.

## 6.3 Advertising spend and sales

```{r advertising}
plot(
  sales_data$Advertising_Spend,
  sales_data$Sales,
  main = "Advertising Spend vs Sales",
  xlab = "Advertising Spend",
  ylab = "Sales",
  pch = 19
)

abline(lm(Sales ~ Advertising_Spend, data = sales_data))
```

**Interpretation:** The scatter plot helps visualize the association between
advertising expenditure and sales.

## 6.4 Distribution of monthly sales

```{r histogram}
hist(
  sales_data$Sales,
  main = "Distribution of Monthly Sales",
  xlab = "Sales",
  ylab = "Number of Months"
)
```

**Interpretation:** The histogram shows the distribution of monthly sales values.

## 6.5 Monthly returns

```{r returns}
plot(
  sales_data$Month,
  sales_data$Returns,
  type = "o",
  main = "Monthly Returns",
  xlab = "Month",
  ylab = "Returns",
  las = 2
)
```

**Interpretation:** This graph shows how the number of returned orders changes over time.

# 7. Findings

1. The dataset contains 24 monthly observations.
2. Total sales during the study period are approximately 1,415,800.
3. Average monthly sales are approximately 58,991.67.
4. The median monthly sales are 58,400.00.
5. Average monthly customers are approximately 693.33.
6. The correlation between advertising spend and sales is approximately 0.999.
7. The correlation between customer count and sales is approximately 1.000.
8. Total recorded returns are 1051.

# 8. Conclusion

The project demonstrates a complete R data-analysis workflow. Monthly sales,
customer count, advertising expenditure and returns were imported, checked,
analysed and visualized. The dataset shows an upward sales trend over the study
period, together with increasing customer counts. Advertising spend and sales
also show a strong positive association in this sample. Correlation describes
association and does not by itself prove that advertising caused the increase
in sales.

# 9. R Code Summary

The project uses `read.csv()` for importing data, `str()`, `is.na()` and
`duplicated()` for preprocessing, `mean()`, `median()`, `sd()` and `sum()` for
statistical analysis, `cor()` for correlation, and `plot()`, `barplot()`,
`hist()` and `lm()` for visualization and trend analysis.

# 10. Final Outcome

The R Markdown file can be opened in RStudio and knitted to HTML. The CSV file
is included in the same folder, so the project does not depend on a
computer-specific file path.
